Data exploration, quality control and statistical analysis of ChIP-exo experiments

### 2015-11-30 rmarkdown/slides.Rnw
added diagrams to slides

\setlength\tabcolsep{3pt}  % default value: 6pt
\scriptsize  %%  command to change the font size
% \rowcolor{RoyalBlue!20} %-- this indicates the change in odd and pair rows
\begin{table}[H]
\begin{tabularx}{\textwidth}{ @{} c|c|c|c|c|c|c| Y @{}}
  \toprule
\textbf{IP} & \textbf{Organism}  & \textbf{Condition} & \textbf{Rep.} & \textbf{Nr. reads} & \textbf{PBC} &  \textbf{NSC} & \textbf{Source} \\
\midrule
$\sig$ & E.Coli & Rif-0min & 1 & 960,256 & 0.2823 &   10.29   & \multirow{4}{\hsize}{Courtesy of Prof. Landick's lab} \\
$\sig$ &  E.Coli & Rif-0min & 2 & 2,247,295 & 0.2656 &  25.08  \\
$\sig$ &  E.Coli & Rif-20min & 1 & 1,940,387 & 0.2698 &  17.69  \\
$\sig$ &  E.Coli & Rif-20min & 2 & 4,229,574 & 0.2153 &   14.11 \\
\hline
FoxA1 &  Mouse &  - & 1 & 22,210,461 & 0.6562 & 21.452 & \multirow{3}{\hsize}{From Serandour et al., 2013 \cite{exoillumina}} \\
FoxA1 &  Mouse &  - & 2 & 22,307,557 & 0.7996 & 60.661 \\
FoxA1 &  Mouse &  - & 3 & 22,421,729 & 0.1068 & 72.312 \\
\hline
ER & Human & - & 1 & 9,289,835 & 0.8082 & 19.843 & \multirow{3}{\hsize}{From Serandour et al., 2013 \cite{exoillumina}} \\
ER & Human & - & 2 & 11,041,833 & 0.8024 & 21.422 \\
ER & Human & - & 3 & 12,464,836 & 0.8203 & 19.699 \\
\hline
CTCF & Human & - & 1 &   48,478,450 & 0.4579 & 15.977 & From Rhee and Pugh 2011, \cite{exo1} \\
\bottomrule
\end{tabularx}  
\end{table}

@@

\begin{frame}
  \frametitle{ARC vs. URCR exploration by replicate}
\only<1>{
\begin{figure}[H]
  \centering
  \includegraphics[width = \textwidth,page = 1]{../figs/Carroll_mice_for_paper/FoxA1_enrichment.pdf}
\end{figure}
}

\only<2>{
\begin{figure}[H]
  \centering
  \includegraphics[width = \textwidth,page = 3]{../figs/Carroll_mice_for_paper/FoxA1_enrichment.pdf}
\end{figure}
}

@@

\only<1>{
\begin{figure}[H]
  \centering
  \includegraphics[width = \textwidth,page = 1]{../figs/for_paper/local_SCC_separated.pdf}
\end{figure}
}

\only<2>{
\begin{figure}[H]
  \centering
  \includegraphics[width = \textwidth,page = 2]{../figs/for_paper/local_SCC_separated.pdf}
\end{figure}
}

\only<3>{
\begin{align}
 y(\delta) = f(x_\delta) + \epsilon_\delta, \quad 
\mbox{local-NSC} = \frac{max_{x_\delta} \hat{f}(x_\delta)}{\hat{\sigma}_f},
\nonumber
\end{align}
\vspace*{\fill}
}

\only<3>{
\begin{columns}

@@

\only<4>{
  \begin{figure}[H]
    \centering
      \includegraphics[width = \textwidth,page =4]{../figs/Carroll_mice_for_paper/Local_SCC_indicator_by_strata.pdf}
  \end{figure}

@@

{\scriptsize
\begin{itemize}
\item  
\end{itemize}
}
\end{frame}

@@

\begin{frame}
  \frametitle{Structure of the experiment}

  \begin{figure}[H]
    \centering

  \end{figure}

@@

\only<2>{
\begin{figure}[H]
\centering
   \includegraphics[width = .46\textwidth,page = 1]{../figs/for_paper/sigma_delta_old_densities.pdf}
   \includegraphics[width = .46\textwidth,page = 2]{../figs/for_paper/sigma_delta_old_densities.pdf} 
 \end{figure}
}
\only<2>{
{\small
\begin{itemize}
\item $\delta$ measures the average distance of reads to their
  respective binding sites
\item $\sigma$ measures the dispersion of reads around their
  respective binding sites
\end{itemize}
}}
\end{frame}

@@

\only<1>{
\begin{figure}[H]
  \centering
  \includegraphics[width = .4\textwidth,page = 1]{../figs/for_paper/saturation_analysis_old_frames.pdf}
  \includegraphics[width = .4\textwidth,page = 2]{../figs/for_paper/saturation_analysis_old_frames.pdf}
\end{figure}
}
\only<1>{
{\small
  \begin{itemize}
  \item ChIP-exo and PET ChIP-Seq are comparable and outperform SET
    ChIP-Seq
  \end{itemize}
}}

@@

@@

  \begin{itemize}
  \item Our pipeline is capable of assessing the balance between
    sample enrichment and library complexity.
  \item We shown that the ``peak-pair'' assumption doesn't hold well
    in practice, and implemented a visualization capable of detecting
    strand imbalance.
  \item We updated dPeak, which makes a striking balance in
    sensitivity, specificity and spatial resolution.
  \item ChIP-exo and PET ChIP-Seq are comparable in resolution and
    sensitivity, and both outperform SET ChIP-Seq.
  \item We showed that with a fixed number of reads, ChIP-exo
    outperforms PET and SET ChIP-Seq.
  \end{itemize}
\end{frame}

@@

\begin{itemize}
\item In the paper, we showed that there is a relationship between
  ChIP-exo tag counts and both mappability and GC content scores. We
  want to add a QC measure to the pipeline based on them.
\item We want to assess if ChIP-Nexus library complexity is actually higher than ChIP-exo's by using the $\mbox{local-NSC}$.
\item We have been studying E. Coli's transcription initiation
  complexes with PET ChIP-Seq, being able to label regions as open or
  closed complexes. We want to improve this analysis by using ChIP-exo
  data, and hopefully detecting intermediate step between this two
  states.
\item Find a optimal strategy for labeling enhancer out of a
  predetermined list of regions in the genome by the use of active
  learning techniques.
\end{itemize}

>>>>

\setlength\tabcolsep{3pt}  % default value: 6pt
\scriptsize  %%  command to change the font size
% \rowcolor{RoyalBlue!20} %-- this indicates the change in odd and pair rows
\begin{table}[H]
\begin{tabularx}{\textwidth}{ @{} c|c|c|c|c|c|c| Y @{}}
  \toprule
\textbf{IP} & \textbf{Organism}  & \textbf{Condition} & \textbf{Rep.} & \textbf{Nr. reads} & \textbf{PBC} &  \textbf{NSC} & \textbf{Source} \\
\midrule
$\sig$ & E.Coli & Rif-0min & 1 & 960,256 & 0.2823 &   10.29   & \multirow{4}{\hsize}{Courtesy of Prof. Landick's lab} \\
$\sig$ &  E.Coli & Rif-0min & 2 & 2,247,295 & 0.2656 &  25.08  \\
$\sig$ &  E.Coli & Rif-20min & 1 & 1,940,387 & 0.2698 &  17.69  \\
$\sig$ &  E.Coli & Rif-20min & 2 & 4,229,574 & 0.2153 &   14.11 \\
\hline
FoxA1 &  Mouse &  - & 1 & 22,210,461 & 0.6562 & 21.452 & \multirow{3}{\hsize}{From Serandour et al., 2013 \cite{exoillumina}} \\
FoxA1 &  Mouse &  - & 2 & 22,307,557 & 0.7996 & 60.661 \\
FoxA1 &  Mouse &  - & 3 & 22,421,729 & 0.1068 & 72.312 \\
\hline
ER & Human & - & 1 & 9,289,835 & 0.8082 & 19.843 & \multirow{3}{\hsize}{From Serandour et al., 2013 \cite{exoillumina}} \\
ER & Human & - & 2 & 11,041,833 & 0.8024 & 21.422 \\
R & Human & - & 3 & 12,464,836 & 0.8203 & 19.699 \\
\hline
CTCF & Human & - & 1 &   48,478,450 & 0.4579 & 15.977 & From Rhee and Pugh 2011, \cite{exo1} \\
\bottomrule
\end{tabularx}  
\end{table}

@@

\begin{frame}
  \frametitle{ARC vs. URCR exploration by replicate}
\only<1>{
All regions
\begin{figure}[H]
  \centering
  \includegraphics[width = \textwidth,page = 1]{../figs/Carroll_mice_for_paper/FoxA1_enrichment.pdf}
\end{figure}
}

\only<2>{
Regions formed by reads aligned to more than 10 unique positions
\begin{figure}[H]
  \centering
  \includegraphics[width = \textwidth,page = 3]{../figs/Carroll_mice_for_paper/FoxA1_enrichment.pdf}
\end{figure}
}

\only<3>{
Regions formed by reads aligned to more than 30 unique positions
\begin{figure}[H]
  \centering
  \includegraphics[width = \textwidth,page = 4]{../figs/Carroll_mice_for_paper/FoxA1_enrichment.pdf}
\end{figure}
}

@@

\only<1>{
Calculate coverage for all regions
\begin{figure}[H]
  \centering
  \includegraphics[width = \textwidth,page = 1]{../figs/for_paper/local_SCC_separated.pdf}
\end{figure}
}

\only<2>{
Calculate $\mbox{local-SCC}$ for regions
\begin{figure}[H]
  \centering
  \includegraphics[width = \textwidth,page = 2]{../figs/for_paper/local_SCC_separated.pdf}
\end{figure}
}

\only<3>{
Fit local polynomial regression for the $\mbox{local-SCC}$ 
\begin{figure}[H]
  \centering
  \includegraphics[width = \textwidth,page = 3]{../figs/for_paper/local_SCC_separated.pdf}
\end{figure}
}

\only<4>{
\begin{align}
 y(\delta) = f(x_\delta) + \epsilon_\delta, \quad 
\mbox{local-NSC} = \frac{max_{x_\delta} \hat{f}(x_\delta)}{\hat{\sigma}_f},
\nonumber
\end{align}
\vspace*{\fill}
}

\only<4>{
\begin{columns}

@@

\only<5>{
  \begin{figure}[H]
    \centering
      \includegraphics[width = \textwidth,page =4]{../figs/Carroll_mice_for_paper/Local_SCC_indicator_by_strata.pdf}
  \end{figure}

@@

{\scriptsize
\begin{itemize}
\item 
\end{itemize}
}

@@

\begin{frame}[plain]
  \frametitle{Data structure}
Data available for $\sig$:
  \begin{figure}[H]
    \centering
    \includegraphics[width = \textwidth]{../figs/for_paper/chip_explanation.pdf}
  \end{figure}

@@

\only<2>{
\begin{figure}[H]
\centering
   \includegraphics[width = .46\textwidth,page = 1]{../figs/for_paper/sigma_delta_old_densities.pdf}
   \includegraphics[width = .46\textwidth,page = 2]{../figs/for_paper/sigma_delta_old_densities.pdf} 
 \end{figure}
}
\only<2>{
{\small
\begin{itemize}
\item $\delta$ measures the average distance of reads to their
  respective binding sites.
\item $\sigma$ measures the dispersion of reads around their
  respective binding sites.
\item Good news !! the model is reflecting the truth.
\end{itemize}
}}
\end{frame}

@@

\only<1>{
\begin{figure}[H]
  \centering
  \includegraphics[width = .4\textwidth,page = 1]{../figs/for_paper/saturation_analysis_old_frames.pdf}
  \includegraphics[width = .4\textwidth,page = 2]{../figs/for_paper/saturation_analysis_old_frames.pdf}
\end{figure}
}

@@

\only<2>{
{\small
  \begin{itemize}
  \item ChIP-exo and PET ChIP-Seq are comparable and outperform SET
    ChIP-Seq
  \end{itemize}
}}

@@

  \begin{itemize}
  \item Our pipeline is capable of assessing the balance between
    sample enrichment and library complexity.
  \item We shown that the ``peak-pair'' assumption doesn't hold well
    in practice, and implemented a visualization capable of detecting
    strand imbalance.
  \item We updated dPeak, which makes a striking balance in
    sensitivity, specificity and spatial resolution.
  \item ChIP-exo and PET ChIP-Seq are comparable in resolution and
    sensitivity, and both outperform SET ChIP-Seq.
  \item We showed that with a fixed number of reads, ChIP-exo
    outperforms PET and SET ChIP-Seq.
  \item dPeak outperforms other algorithms in resolution.
  \end{itemize}
\end{frame}

@@

\begin{itemize}
\item In the paper, we showed that there is a relationship between
  ChIP-exo tag counts and both mappability and GC content scores. We
  want to add a QC measure to the pipeline based on them.
\item We want to assess if ChIP-Nexus library complexity is actually
  higher than ChIP-exo's by using the $\mbox{local-NSC}$.
\item We have been studying E. Coli's transcription initiation
  complexes with PET ChIP-Seq. We want to improve this analysis by
  using ChIP-exo data.
\item Find a optimal strategy for labeling enhancer out of a
  predetermined list of regions in the genome by the use of active
  learning techniques.
\end{itemize}
